---
title: "20210130-millRiver_index_jsg2145"
author: "Jared Garfinkel"
date: "1/30/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(readxl)
library(corrplot)
library(factoextra)
library(gridExtra)
library(psych)
library(caret)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 12, 
  fig.height = 14,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# read in data

```{r}
df = readxl::read_excel("./data/Index_Dataset_01-26-21 (1).xlsx") %>% 
  janitor::clean_names() %>% 
  select(overall_resilience, everything())

skimr::skim(df)
```

This dataset is `r nrow(data)` rows by `r ncol(data)` columns. There are 35 indicators including 5 summary statistics, "overall_resilience", "health", "diversity", "social", and "environmental". These summary statistics are removed for analysis. 

There is no missing data.

```{r}
data = df %>% 
  select(-overall_resilience, -health, -economic, -social, -environmental)
```

# Exploratory Data Analysis

## conduct a correlation analysis

```{r, fig.height = 12, fig.width = 10}
M = cor(data[,-1])
# ?corrplot
pdf("corrplot_output.pdf")
corrplot(M, type = c("upper"), order = c("hclust"))
dev.off()
```

There are several race/ethnicity variables that are correlated with each other and economic- and education-related variables.

It will be easier to examine highly correlated variables using a cutoff to view those correlations.

We start with a cutoff of Pearson's R-squared $\ge$ 0.7, which results in the following:

```{r}
output = vector(mode = "list", length = ncol(M))
rname = NULL
cname = NULL
for(i in 1:nrow(M)) {
  for(j in 1:ncol(M)) {
    rname[[i]] = dimnames(M)[[c(1, i)]]
    cname[[j]] = dimnames(M)[[c(2, j)]]
    output[[c(i, j)]] = tibble_row(i, j, rname[[i]], cname[[j]], M[[i, j]])
  }
}
```

```{r, result = "hide"}
matrify = function(pdata = output) {
  for(i in 1:ncol(M)) {
      result[[i]] = tibble(pdata[[c(i, i)]])
  }
  result = bind_rows(result)
  return(result)
}
```

# Matrify 2

```{r}
matrify2 = function(pdata = output) {
  result = NULL
  for(i in 1:length(pdata)) {
      result[[i]] = tibble(pdata[[i]])
  }
  result = bind_rows(result)
  return(result)
}

df2 = vector(mode = "list", length = ncol(M))
for(j in 1:ncol(M)) {
  df2[[j]] = matrify2(pdata = output[[j]])
}

# head(df2)
```

```{r}
union_df = bind_rows(df2) %>% 
  rename("r2_value" = "M[[i, j]]",
         "column" = "cname[[j]]",
         "row" = "rname[[i]]")
```

```{r}
head(union_df)
nrow(union_df)
```



```{r}
high_df = union_df %>% 
  filter(abs(r2_value) >= 0.7,
         r2_value != 1,
         i < j)

pdf("high_corr.pdf")
high_df %>% 
  select(row, column, r2_value) %>%
  rename("Row" = row, "Column" = column, "Pearson-R2" = r2_value) %>% 
  knitr::kable()
dev.off()
```


```{r}
very_high_df = union_df %>% 
  filter(abs(r2_value) >= 0.8,
         r2_value != 1,
         i < j)

very_high_df
```

The indicators teen_pregnancy and asthma are unexpectedly very highly correlated.

# Unsupervised learning

```{r}
csdf = scale(data[,-1])
rownames(csdf) = data$community
```


## Principal Component Analysis

### prcomp

```{r}
pca1 = prcomp(csdf)
fviz_eig(pca1)
```

```{r}
a = fviz_contrib(pca1, choice = "var", axes = 1) 
b = fviz_contrib(pca1, choice = "var", axes = 2) 
grid.arrange(a, b, nrow = 2)
```

```{r}
biplot(pca1)
```


### fa.parallel

```{r}
set.seed(719)
fa.parallel(csdf) # parallel analysis suggests the number of factors = 4, and the number of components = 3
```

### principal

```{r}
fit = principal(M, nfactors = 4, rotate = "varimax")
fit
```

### iclust

```{r, fig.height = 16, fig.width = 14}
# ?iclust
pdf("iclust_graph.pdf")
fit2 = iclust(M, plot = TRUE)
dev.off()
fit2
```

## Clustering

Further  methods in unsupervised learning

```{r}
# ?fviz_nbclust
fviz_nbclust(csdf,
             FUNcluster = kmeans,
             method = "silhouette")
```

```{r}
set.seed(719)
km <- kmeans(csdf, centers = 2)
```

```{r}
output2 = NULL
for(i in 1:length(km$cluster)) {
  output2[[i]] = tibble(km$cluster[[i]])
}

output2_df = bind_rows(output2) %>% 
  rename("cluster" = "km$cluster[[i]]")
```

```{r}
pdf("2cluster_table.pdf")
data %>% 
  add_column(cluster = pull(output2_df, cluster)) %>% 
  select(cluster, community) %>% 
  arrange(cluster) %>% 
  rename("Cluster" = cluster, "Community" = community) %>% 
  knitr::kable()
dev.off()
```

```{r, fig.height = 12, fig.width = 14}
# ?fviz_cluster
km_vis <- fviz_cluster(list(data = csdf, cluster = km$cluster),
                       ellipse.type = "convex", 
                       geom = c("point","text"),
                       labelsize = 10) + 
  labs(title = "K-means") 

# pdf("2cluster_graph.pdf")
km_vis
# dev.off()
```

# Hierarchical Clustering

```{r}
hc.complete <- hclust(dist(csdf), method = "complete")
set.seed(719)
fviz_dend(hc.complete, k = 2,        
          cex = 0.5, 
          palette = "jco", 
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, rect_border = "jco",
          labels_track_height = 2.5)

# ?fviz_dend
pdf("cluster_dendrogram.pdf")
ind4.complete <- cutree(hc.complete, 2)
dev.off()
```

# variance importance in caret

```{r caret setup}
ctrl1 = trainControl(method = "repeatedcv", number = 10, repeats = 5)

# ?train

df2 = df %>% 
  select(-environmental, -social, -health, -economic) %>% 
  data.frame()

rownames(df2) = df2$community

df2 = df2 %>% 
  select(-community)
```


```{r glm in caret}
mod1 = train(x = df2[,-1],
             y = df2[,1],
                method = "glm",
                preProcess = c("center","scale"),
                trControl = ctrl1)

varimp1 = varImp(mod1)

pdf("glm_varimp.pdf")
varimp1$importance %>% 
  arrange(desc(Overall)) %>% 
  knitr::kable()
dev.off()
```

```{r gam in caret}
set.seed(719)
mod2 = train(x = df2[,-1],
             y = df2[,1],
                method = "earth",
                preProcess = c("center","scale"),
                trControl = ctrl1)

varimp2 = varImp(mod2)

pdf("gam_varimp.pdf")
varimp2$importance %>% 
  arrange(desc(Overall)) %>% 
  knitr::kable()
dev.off()
```



